Attrition in Online Learning: Understanding Persistence and Dropout in Massive Open Online Courses

### 2014-10-27 paper_v1.Rnw
Testing pushing changes to .Rnw only

\begin{abstract}
Persistence and attrition in MOOCs are systematically analyzed using self-report and behavioral data collected from [N] online learners in [M] courses. Study 1 offers insights into reasons for disengaging from MOOCs and explores relationships with prior behavior and reported intentions. Study 2 is a case study to develop a deeper understanding of attrition in MOOCs by conducting a case study. Targeting online learners who were predicted likely dropouts in a particular course were invited to provide feedback via a survey. (abstract needs work...)
\end{abstract}

@@

TODO: Outline different courses and go through variabels in the course info table.

Table:  course name; enrollment; pre-survey response rate; prop. demographics (age, gender, educ); who intended to do all; post-survey response rate; self-identified dropout rate

@@

Table: proportions for each dropout reason for each course

>>>>

\begin{abstract}
Persistence and attrition in MOOCs are systematically analyzed using self-report and behavioral data collected from 20879 online learners in 18 courses. Study 1 offers insights into reasons for disengaging from MOOCs and explores relationships with prior behavior and reported intentions. Study 2 is a case study to develop a deeper understanding of attrition in MOOCs by conducting a case study. Targeting online learners who were predicted likely dropouts in a particular course were invited to provide feedback via a survey. (abstract needs work...)
\end{abstract}

@@

% Which courses did we use? What data
In this study, we used 17 MOOCs from different areas including humanities sciences (4 courses), engineering (2 courses), computer science (3 courses), mathematics and statistics (6 courses), and business (2 instances of 1 course). For each course, we acquired the data from the pre-course and post-course surveys, as well as the learner activity data from the platform hosting the course. From each course, we selected all learners who took the post-course survey, and we matched learners to their activity data and pre-course survey response (if available). Table \ref{tab:study1_courses_summary} shows the number of enrolled learners, number of learners who responses to the pre-course/post-course survey, distributions of age, gender, and education level, and the proportion of learner who self-reported that they stopped participating in the course. We also computed the proprtions of learners who viewed at least $p$\% of the course videos for $p = 30\%$, $50\%$, and $80\%$.

%Table:  course name; enrollment; pre-survey response rate; prop. demographics (age, gender, educ); who intended to do all; post-survey response rate; self-identified dropout rate
\newcommand{\specialcellcentered}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
\newcommand{\specialcellleft}[2][c]{%
  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}

\begin{table*}
\label{tab:study1_courses_summary}
\caption{Summary on courses and surveys used in Study 1}
\small
\center
\begin{tabular}{llrrrrrrrrrr}
\toprule
ID & Topic Area & Enrollment & Pre-survey & Post-survey & Female & Age & Education & \specialcellcentered{\ensuremath{>} 30\% \\ videos} & \specialcellcentered{\ensuremath{>} 50\% \\ videos} & \specialcellcentered{\ensuremath{>} 80\% \\ videos} & \specialcellcentered{Self-Rep. \\ Dropout} \\
\midrule
C1 & Business & 61096 & 9943 & 2367 & 42\% & 39 (30,49) & n/a & 39\% & 20\% & 7\% & 10\% \\
C2 & Business & 41081 & 6544 & 883 & 40\% & 39 (30,49) & 97\% & 32\% & 13\% & 2\% & 9\% \\
C3 & Comp. Sc. & 27104 & 6104 & 2854 & 35\% & 30 (22,41) & 74\% & 67\% & 40\% & 7\% & 2\% \\
C4 & Comp. Sc. & 56154 & 3406 & 1123 & 11\% & 33 (27,42) & 91\% & 34\% & 16\% & 11\% & 21\% \\
C5 & Comp. Sc. & 57636 & 5738 & 402 & 17\% & 33 (26,43) & 91\% & 66\% & 55\% & 28\% & 14\% \\
C6 & Engineering & 16220 & 2540 & 269 & 13\% & 33 (27,45) & 92\% & 62\% & 49\% & 17\% & 15\% \\
C7 & Enginerring & 7178 & 1312 & 202 & 12\% & 27 (22,39) & 85\% & 51\% & 18\% & 2\% & 7\% \\
C8 & Humanities & 7634 & 3631 & 152 & 92\% & 33 (26,45) & 93\% & 45\% & 11\% & 0\% & 7\% \\
C9 & Humanities & 6475 & 3315 & 326 & 90\% & 32 (27,47) & 91\% & 69\% & 50\% & 17\% & 16\% \\
C10 & Humanities & 9903 & 4227 & 1570 & 52\% & 39 (29,52) & 89\% & 84\% & 74\% & 21\% & 3\% \\
C11 & Humanities & 630 & 167 & 80 & 50\% & 30 (27,40) & 97\% & 74\% & 55\% & 16\% & 16\% \\
C12 & Math \& Stats & 49317 & 3090 & 862 & 33\% & 41 (29,55) & 88\% & 27\% & 10\% & 1\% & 31\% \\
C13 & Math \& Stats & 32006 & 26142 & 5336 & 58\% & 28 (16,46) & n/a & 69\% & 38\% & 8\% & 4\% \\
C14 & Math \& Stats & 16857 & 2892 & 363 & 9\% & 31 (26,38) & 99\% & 61\% & 45\% & 25\% & 14\% \\
C15 & Math \& Stats & 37017 & 11629 & 1854 & 20\% & 35 (29,45) & 98\% & 78\% & 67\% & 34\% & 8\% \\
C16 & Math \& Stats & 27236 & 2711 & 637 & 82\% & 43 (34,51) & n/a & 64\% & 36\% & 5\% & 1\% \\
C17 & Math \& Stats & 13132 & 3629 & 843 & n/a & n/a & n/a & 70\% & 54\% & 14\% & 5\% \\
\bottomrule
\end{tabular}
\end{table*}

% What LMS data variables did we compute?
We are interested in exploring correlations between reasons for dropout and learner behaviors. If such correlations exist, they can be an additional towards understanding why learners engage with or disengage from various learning activities. In this study, we computed the fraction of videos watched by each learner, fraction of problems attempted, average grade scored on attempted problems, and number of forum posts and replies posted.

@@

%Table: proportions for each dropout reason for each course
\begin{table*}
\label{tab:proprtion_reason_for_dropout}
\caption{Distribution of reported reasons for dropping out across courses. Percentage of learners who selected each reason for dropping out in each course, and median and interquartile range across courses}
\small
\center
\begin{tabular}{llrrrrrrrrrrrrrrrrrr}
\toprule
Reason for dropout & C1 & C2 & C3 & C4 & C5 & C6 & C7 & C8 & C9 & C10 & C11 & C12 & C13 & C14 & C15 & C16 & C17 & Median & IQR \\
\midrule
\specialcellleft{Material not explained \\ properly} & 18 & 20 & 10 & 22 & 19 & 21 & 8 & 11 & 13 & 13 & 9 & 7 & 14 & 18 & 11 & 14 & 19 & 14 & 8 \\
\specialcellleft{Course requires too \\ much time} & 57 & 39 & 13 & 48 & 42 & 51 & 33 & 60 & 65 & 42 & 18 & 52 & 21 & 68 & 39 & 28 & 68 & 42 & 24 \\
\specialcellleft{Not what I was looking \\ for} & 21 & 23 & 25 & 14 & 13 & 19 & 8 & 0 & 0 & 23 & 9 & 12 & 20 & 20 & 13 & 28 & 12 & 14 & 9 \\
I joined too late & 32 & 18 & 25 & 28 & 32 & 28 & 25 & 11 & 20 & 27 & 27 & 19 & 19 & 22 & 18 & 0 & 11 & 22 & 9 \\
\specialcellleft{Didn't enjoy online \\ course format} & 15 & 14 & 8 & 16 & 13 & 12 & 0 & 0 & 9 & 10 & 9 & 9 & 19 & 12 & 7 & 14 & 14 & 12 & 5 \\
Technical difficulties & 21 & 3 & 18 & 12 & 17 & 24 & 25 & 30 & 30 & 10 & 36 & 13 & 20 & 25 & 12 & 0 & 21 & 20 & 13 \\
\specialcellleft{Couldn't keep up with \\ deadlines} & 69 & 58 & 38 & 76 & 80 & 66 & 66 & 55 & 69 & 67 & 27 & 71 & 28 & 64 & 54 & 0 & 71 & 66 & 15 \\
Not challenging enough & 12 & 7 & 27 & 8 & 4 & 18 & 41 & 11 & 6 & 10 & 27 & 10 & 24 & 14 & 11 & 28 & 7 & 11 & 16 \\
I learned all I wanted & 18 & 7 & 26 & 17 & 19 & 16 & 41 & 10 & 9 & 25 & 18 & 21 & 27 & 25 & 20 & 28 & 12 & 19 & 9 \\
Videos were confusing & 13 & 12 & 10 & 19 & 19 & 21 & 0 & 11 & 7 & 13 & 9 & 9 & 11 & 16 & 9 & 14 & 19 & 12 & 7 \\
I was only exploring & 16 & 5 & 22 & 17 & 14 & 21 & 16 & 11 & 7 & 18 & 20 & 16 & 23 & 20 & 16 & 16 & 12 & 16 & 6 \\
\specialcellleft{Too advanced for my \\ background} & 14 & 10 & 10 & 23 & 24 & 16 & 8 & 11 & 2 & 13 & 18 & 19 & 10 & 25 & 17 & 0 & 29 & 14 & 9 \\
\specialcellleft{Assn or exam was too \\ difficult} & 26 & 8 & 11 & 27 & 23 & 39 & 33 & 11 & 24 & 11 & 27 & 27 & 11 & 35 & 15 & 14 & 41 & 24 & 16 \\
\specialcellleft{Credentials not valuable \\ to me} & 14 & 9 & 18 & 10 & 10 & 18 & 25 & 22 & 4 & 5 & 18 & 7 & 20 & 22 & 14 & 14 & 25 & 14 & 10 \\
\bottomrule
\end{tabular}
\end{table*}
